Human capital drives natural selection in contemporary humans

### 2021-04-14 model-appendix.Rmd
responses to PNAS review
Address comments in PNAS review - new title - Use standard_ggplot more - Add confidence intervals - Main plot to start - Discuss endogeneity - Add computation of CARA utility

\externaldocument{why-natural-selection}

\section*{Appendix}

@@

Rearranging (\ref{eq:kkt1}), (\ref{eq:kkt2}) and (\ref{eq:kkt3})
gives:
\begin{align}
N_{1}^{*} & =\frac{1}{b}\left(1-s^{*}-\left(\frac{b}{a}\right)^{1/\sigma}\right);\label{eq:N11}\\
N_{2}^{*} & =\frac{1}{b}\left(1-\left(\frac{b}{a}\right)^{1/\sigma}(s^{*}h)^{(1-\sigma)/\sigma}\right);\label{eq:N21}\\
s^{*} & =\frac{1-bN_{1}^{*}}{1+\left((1-bN_{2}^{*})h\right)^{1-1/\sigma}}.
\end{align}
Plugging the expressions for $N_{1}^{*}$ and $N_{2}^{*}$ into $s^{*}$gives
\[
s^{*}=\frac{s^{*}+\left(\frac{b}{a}\right)^{1/\sigma}}{1+\left(\left(\frac{b}{a}\right)^{1/\sigma}s^{*(1-\sigma)/\sigma}h{}^{1/\sigma}\right)^{1-1/\sigma}}
\]
which simplifies to
\begin{equation}
s^{*}=\left(\frac{b}{a}\right)^{1/(2\sigma-1)}h^{(1-\sigma)/(2\sigma-1)}.\label{eq:s1}
\end{equation}

@@

From Lemmas \ref{case-1-lemma}, \ref{case-2-lemma} and \ref{case-3-lemma},
as $h$ increases we move from $N_{1}^{*},N_{2}^{*}>0$ to $N_{1}^{*}=0,N_{2}^{*}>0$
to $N_{1}^{*}=N_{2}^{*}=0$. Furthermore, for $\sigma>2/3$, $N_{1}^{*}$
and $N_{2}^{*}$ are convex in $h$ when they are both positive, and
for $\sigma$ close enough to 1, $N_{2}^{*}$ is convex in $h$ when
$N_{1}^{*}=0$. To show global concavity of $N^{*}$ in $h$ as $\sigma\nearrow1$,
all that remains is to check that the derivative is increasing around
the points where these 3 regions meet. That is trivially satisfied
where $N_{2}^{*}$ becomes 0, since thereafter $\frac{dN^{*}}{dh}$
is zero. The derivative as $N_{1}^{*}$ approaches zero is twice the
expression in (\ref{eq:dNdh1}):
\begin{equation}
-\frac{2}{b}\left(\frac{b}{a}\right)^{1/(2\sigma-1)}\frac{1-\sigma}{2\sigma-1}h^{(1-\sigma)/(2\sigma-1)-1}\label{eq:dNonleft}
\end{equation}
and the derivative to the right of this point is given by (\ref{eq:dNdh2}):
\begin{equation}
-\frac{1}{b}\left(\frac{b}{a}\right)^{1/\sigma}\frac{1-\sigma}{\sigma}(s^{*}h)^{(1-2\sigma)/\sigma}(s^{*}+h\frac{ds^{*}}{dh})\label{eq:dNonright}
\end{equation}

@@

Examining (\ref{eq:estar2}) shows that as $\sigma\rightarrow1$,
$s^{*}\rightarrow0.5$ and $\frac{ds^{*}}{dh}\rightarrow0$, and therefore
the above approaches
\[
2\frac{1}{(0.5)^{2}}=8.
\]

Thus, as $\sigma\nearrow1$, $N^{*}$ is guaranteed to be globally
convex in $h$.

We can now prove the facts stated in the main text.

>>>>

@@

Rearranging (\ref{eq:kkt1}), (\ref{eq:kkt2}) and (\ref{eq:kkt3})
gives:
\begin{align}
N_{1}^{*} & =\frac{1}{b}\left(1-s^{*}-\left(\frac{b}{a}\right)^{1/\sigma}\right);\label{eq:N11}\\
N_{2}^{*} & =\frac{1}{b}\left(1-\left(\frac{b}{a}\right)^{1/\sigma}(s^{*}h)^{(1-\sigma)/\sigma}\right);\label{eq:N21}\\
s^{*} & =\frac{1-bN_{1}^{*}}{1+\left((1-bN_{2}^{*})h\right)^{1-1/\sigma}}.
\end{align}
Plugging the expressions for $N_{1}^{*}$ and $N_{2}^{*}$ into $s^{*}$ gives
\[
s^{*}=\frac{s^{*}+\left(\frac{b}{a}\right)^{1/\sigma}}{1+\left(\left(\frac{b}{a}\right)^{1/\sigma}s^{*(1-\sigma)/\sigma}h{}^{1/\sigma}\right)^{1-1/\sigma}}
\]
which simplifies to
\begin{equation}
s^{*}=\left(\frac{b}{a}\right)^{1/(2\sigma-1)}h^{(1-\sigma)/(2\sigma-1)}.\label{eq:s1}
\end{equation}

@@

\begin{lem}
For $\sigma < 1$ in a neighbourhood of 1, $N^*$ is globally convex in $h$.
\end{lem}
\begin{proof}
From Lemmas \ref{case-1-lemma}, \ref{case-2-lemma} and \ref{case-3-lemma},
as $h$ increases we move from $N_{1}^{*},N_{2}^{*}>0$ to $N_{1}^{*}=0,N_{2}^{*}>0$
to $N_{1}^{*}=N_{2}^{*}=0$. Furthermore, for $\sigma>2/3$, $N_{1}^{*}$
and $N_{2}^{*}$ are convex in $h$ when they are both positive, and
for $\sigma$ close enough to 1, $N_{2}^{*}$ is convex in $h$ when
$N_{1}^{*}=0$. All that remains is to check that the derivative is increasing around
the points where these 3 regions meet. That is trivially satisfied
where $N_{2}^{*}$ becomes 0, since thereafter $\frac{dN^{*}}{dh}$
is zero. The derivative as $N_{1}^{*}$ approaches zero is twice the
expression in (\ref{eq:dNdh1}):
\begin{equation}
-\frac{2}{b}\left(\frac{b}{a}\right)^{1/(2\sigma-1)}\frac{1-\sigma}{2\sigma-1}h^{(1-\sigma)/(2\sigma-1)-1}\label{eq:dNonleft}
\end{equation}
and the derivative to the right of this point is given by (\ref{eq:dNdh2}):
\begin{equation}
-\frac{1}{b}\left(\frac{b}{a}\right)^{1/\sigma}\frac{1-\sigma}{\sigma}(s^{*}h)^{(1-2\sigma)/\sigma}(s^{*}+h\frac{ds^{*}}{dh})\label{eq:dNonright}
\end{equation}

@@

Examining (\ref{eq:estar2}) shows that as $\sigma\rightarrow1$,
$s^{*}\rightarrow0.5$ and $\frac{ds^{*}}{dh}\rightarrow0$, and therefore
the above approaches
\[
2\frac{1}{(0.5)^{2}}=8.
\]
\end{proof}

We can now gather the facts stated in the main text.
